Predictions of Occupied Beds at Faculties Providing Daily Shelter And/Or Overnight Services Using Multivariate Linear Regression Implies That Modelling This Situation Using Other Models Would Be Better

### 2024-12-14 paper/paper.qmd
Make one final change

---
title: "Predictions of Occupied Beds at Faculties Providing Daily Shelter And/Or Overnight Services Using Multivariate Linear Regression Implies That Modelling This Situation Using Other Models Would Be Better"
author: 
  - Kevn Shao
thanks: "Code and data are available at: https://github.com/KevinShao1357/Toronto_Daily_Shelter_And_Overnight_Service_Modelling.git."
date: today
date-format: long
abstract: "Recently, the Canadian government expresses interest in continuous investment in government faculties providing daily shelter and/or overnight services. With the aim to evaluate the validity of predicting the occupied beds at a related faculties at a day, which can reflect the demand of such services at such locations, a multivariate linear regression model is constructed, with the city to Downtown Toronto, and the gender and age set to Mixed Adults. Five predictor variables are used, including two numerical variables, occupancy rate and capacity, and three categorical variables, including classification, service types, and program types. The final results showed likely violations of linearity assumptions such as homoscedasticity and normality of errors, and suggests that we may need to consider either proceed deeper in the linear regression model by adding nonlinear terms for predictor variables, identifying and deleting influential points, and/or consider the weighted-least squares (WLS) method, or consider switching to other more-flexible model types."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
geometry: margin=0.75in
header-includes:
  - \clearpage
---

@@

Canada provides shelters for people in need, such as refugees, the homeless, and people experiencing transitions in living spaces, and the country is continuing to build new shelters for such residents. On April 16, 2024, the Canadian government announced a funding of over 250 million dollars over two years to address homelessness and encampment all over the country. In 2022, Canada announced a target to eliminate homelessness by 2030. Canada Mortgage and Housing Corporation also allocated a total of 420 million dollars to account for homelessness (@Shelter). This implicates that the Canadian government will continuously invest in improving current shelters and building new shelters around the country. In 2024, Canada's Minister of Immigration, Refugees, and Citizenship Marc Miller also announced that the country would accept 27,000 new refguees this year, and this trend is likely to continue in the next years (@Refugee). We can conclude that more shelters are likely to be built in Canada in the future, and the same trend is also valid for the GTA(Greater Toronto Area).

@@

The dataset "Daily Shelter & Overnight Service Occupancy & Capacity" (@Dataset) was sourced from the dataset with same name found in Open Data Toronto. The paper's in-depth analysis and prelimary data analysis are all completed using the statistical programming language R (@R). The book 'Telling Stories with Data' (@tellingstories) was also used to complete this paper.

@@

knitr::kable(c(Emergency = sum(cleaned_data2$classification == 'Emergency'),
               Transitional = sum(cleaned_data2$classification == 'Transitional')),
             align = "c", col.names = ("Counts For Each Classification"))
```
From @tbl-table4, we can see that there are much more emergency locations (around 5400) than transitional locations (around 1400). There are around three times more emergency locations than transitional locations.

@@

knitr::kable(c(Refugee = sum(cleaned_data2$program_area == 'Base Program - Refugee'),
               Base_Shelter = sum(cleaned_data2$program_area == 'Base Shelter and Overnight Services System'),
               Temp_Refugee = sum(cleaned_data2$program_area == 'Temporary Refugee Response'),
               Winter_Programs = sum(cleaned_data2$program_area == 'Winter Programs')),
             align = "c", col.names = ("Counts For Each Classification"))
```
From @tbl-table5, we observe that locations with the program area 'Base Shelter and Overnight Services System' has most observations (around 4800). Locations with the program area 'Base Program - Refugee' has 1272 observations, around one-fourth of the previous one. The remaining two locations has much less observations.

@@

knitr::kable(c(mean = mean(cleaned_data2$occupancy_rate), standard_deviation = sd(cleaned_data2$occupancy_rate), maximum = max(cleaned_data2$occupancy_rate), minimum = min(cleaned_data2$occupancy_rate)), align = "c", col.names = ("Occupancy Rate in a Given Day"))
```
From @tbl-table7, we can conclude that occupancy rate of locations are all pretty much near 100, with a relatively low standard deviation of 5 percentage points, so most locations are relatively full, but we can still consider it as a predictor variable, but needs careful consideration when interpreting results.

@@

The two qq plots have barely any difference, and with the previous analysis of the residuals standard errors and p-values, we still eliminate the original model and choose the model with interaction terms but before we log the response variable. Now we proceed with the Box-Cox transformation, update the dataset and model, and then create the residuals versus fitted (@fig-figure5) and qq plot (@fig-figure6) for this new model.

@@

The qq plot of the new model after completing the box-cox transformation (@fig-figure6) has improved compared to the previous one. The end with the more negative theoretical quantiles values now only has gradual increasing deviations from the ideal line that the data points should follow, which is marked in red. However, a heavy tail still exists at the end with positive theoretical quantiles, suggesting possible outliers when we only look at the qq plot itself.

@@

In this derived multivariate linear regression model, date is not a predictor variable, and the model is based on a dataset that only considers data from January 1 to December 1, 2024. However, one could make it a continuous predictor variable to increase precision and/or accuracy. For example, if the Canadian government decide to invest in such facilities that provide daily shelter and/or overnight service on November 16, 2026, one could make use of the newly-derived function, and put the date as the date predictor variable, and get a rough prediction of how many beds would be occupied at that specific date and location for a specific facility, aiding them to effectively use investments to improve the entire shelter system.

>>>>

---
title: "Predictions of Occupied Beds at Faculties Providing Daily Shelter And/Or Overnight Services Using Multivariate Linear Regression Implies That Modelling This Situation Using Other Models Would Be Better"
author: 
  - Kevn Shao
thanks: "Code and data are available at: https://github.com/KevinShao1357/Toronto_Daily_Shelter_And_Overnight_Service_Modelling.git."
date: today
date-format: long
abstract: "Recently, the Canadian government expresses interest in continuous investment in government faculties providing daily shelter and/or overnight services. With the aim to evaluate the validity of predicting the occupied beds at a related faculties at a day, which can reflect the demand of such services at such locations, a multivariate linear regression model is constructed, with the city to Downtown Toronto, and the gender and age set to Mixed Adults. Five predictor variables are used, including two numerical variables, occupancy rate and capacity, and three categorical variables, including classification, service types, and program types. The final results showed likely violations of linearity assumptions such as homoscedasticity and normality of errors, and suggests that we may need to consider either proceed deeper in the linear regression model by adding nonlinear terms for predictor variables, identifying and deleting influential points, and/or consider the weighted-least squares (WLS) method, or consider switching to other more-flexible model types."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
geometry: margin=0.75in
---

@@

Canada provides shelters for people in need, such as refugees, the homeless, and people experiencing transitions in living spaces, and the country is continuing to build new shelters for such residents. On April 16, 2024, the Canadian government announced a funding of over 250 million dollars over two years to address homelessness and encampment all over the country. In 2022, Canada announced a target to eliminate homelessness by 2030. Canada Mortgage and Housing Corporation also allocated a total of 420 million dollars to account for homelessness (@Shelter). This implicates that the Canadian government will continuously invest in improving current shelters and building new shelters around the country. In 2024, Canada's Minister of Immigration, Refugees, and Citizenship Marc Miller also announced that the country would accept 27,000 new refugees this year, and this trend is likely to continue in the next years (@Refugee). We can conclude that more shelters are likely to be built in Canada in the future, and the same trend is also valid for the GTA(Greater Toronto Area).

@@

The dataset "Daily Shelter & Overnight Service Occupancy & Capacity" (@Dataset) was sourced from the dataset with same name found in Open Data Toronto. Note that the version of the dataset used for this paper is updated to December 13, 2024. The paper's in-depth analysis and preliminary data analysis are all completed using the statistical programming language R (@R). The book 'Telling Stories with Data' (@tellingstories) was also used to complete this paper.

@@

knitr::kable(c(Emergency = sum(cleaned_data2$classification == 'Emergency'),
               Transitional = sum(cleaned_data2$classification == 'Transitional')),
             align = "c", col.names = ("Counts For Each Classification"))
```
From @tbl-table4, we can see that there are much more emergency locations (around 5600) than transitional locations (around 1400). There are around three times more emergency locations than transitional locations.

@@

knitr::kable(c(Refugee = sum(cleaned_data2$program_area == 'Base Program - Refugee'),
               Base_Shelter = sum(cleaned_data2$program_area == 'Base Shelter and Overnight Services System'),
               Temp_Refugee = sum(cleaned_data2$program_area == 'Temporary Refugee Response'),
               Winter_Programs = sum(cleaned_data2$program_area == 'Winter Programs')),
             align = "c", col.names = ("Counts For Each Classification"))
```
From @tbl-table5, we observe that locations with the program area 'Base Shelter and Overnight Services System' has most observations (around 5000). Locations with the program area 'Base Program - Refugee' has 1316 observations, around one-fourth of the previous one. The remaining two locations has much less observations.

@@

knitr::kable(c(mean = mean(cleaned_data2$occupancy_rate), standard_deviation = sd(cleaned_data2$occupancy_rate), maximum = max(cleaned_data2$occupancy_rate), minimum = min(cleaned_data2$occupancy_rate)), align = "c", col.names = ("Occupancy Rate in a Given Day"))
```
From @tbl-table7, we can conclude that occupancy rate of locations are all pretty much near 100, with a relatively low standard deviation of 6 percentage points, so most locations are relatively full, but we can still consider it as a predictor variable, but needs careful consideration when interpreting results.

@@

The two qq plots both have bad performance, both clearly violating normality of errors assumptions, and with the previous analysis of the residuals standard errors and p-values, we still eliminate the original model and choose the model with interaction terms but before we log the response variable. Now we proceed with the Box-Cox transformation, update the dataset and model, and then create the residuals versus fitted (@fig-figure5) and qq plot (@fig-figure6) for this new model.

@@

The qq plot of the new model after completing the box-cox transformation (@fig-figure6) has improved compared to the previous one, with at least most data points following the trend marked by the red line, but heavy tails can still be seen on both ends, the one with the negative theoretical quantiles and the one with the positive theoretical quantiles. This suggests possible outliers on both ends when we only look at the qq plot itself.

@@

In this derived multivariate linear regression model, date is not a predictor variable, and the model is based on a dataset that only considers data from January 1 to December 13, 2024. However, one could make it a continuous predictor variable to increase precision and/or accuracy. For example, if the Canadian government decide to invest in such facilities that provide daily shelter and/or overnight service on November 16, 2026, one could make use of the newly-derived function, and put the date as the date predictor variable, and get a rough prediction of how many beds would be occupied at that specific date and location for a specific facility, aiding them to effectively use investments to improve the entire shelter system.
